
---
title: "Lab #4: Scraping (into) the Hall of Fame"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Maurycy's Work
## Loading packages

```{r}
library(rvest)
library(dplyr)
library(readr)
library(Lahman)
library(ggplot2)
```

## First, let's see what we're trying to match

```{r}
# Looking at the HallOfFame data to see what format we need
head(HallOfFame, 5)
```

## Scraping the 2025 data

```{r}
# Getting the webpage
url <- "https://www.baseball-reference.com/awards/hof_2025.shtml"
page <- read_html(url)

# Grabbing the tables
tables <- html_table(page)
data <- tables[[1]]

# Quick look at what we got
head(data)
```

## Fixing the column names

The first row has the real column names, so let's fix that:

```{r}
# Moving first row to be the column names
colnames(data) <- data[1, ]
data <- data[-1, ]

head(data)
```

## Getting the important stuff

```{r}
# Pulling out just the columns we need
voting_data <- data %>%
  select(Name, Votes, `%vote`) %>%
  mutate(
    Votes = as.numeric(Votes),
    percent = parse_number(`%vote`)
  )

head(voting_data)
```

## Figuring out who got inducted

```{r}
# Finding total ballots and votes needed
# The top person's percentage tells us the total
top_votes <- max(voting_data$Votes, na.rm = TRUE)
top_percent <- max(voting_data$percent, na.rm = TRUE)

total_ballots <- round(top_votes / (top_percent / 100))
votes_needed <- ceiling(total_ballots * 0.75)

cat("Total ballots:", total_ballots, "\n")
cat("Votes needed to get in:", votes_needed, "\n")

# Adding inducted column
voting_data <- voting_data %>%
  mutate(inducted = ifelse(Votes >= votes_needed, "Y", "N"))

# Who got in?
voting_data %>% 
  filter(inducted == "Y") %>%
  select(Name, Votes, percent)
```

## Matching names to playerIDs

This is the tricky part - we need to find the playerID for each person:

```{r}
# Loading player info
data(People)

# Making a simple function to find IDs
get_playerID <- function(name) {
  # Cleaning up the name
  name <- gsub("\\*|†", "", name)
  name <- trimws(name)
  
  # Splitting into first and last name
  parts <- strsplit(name, " ")[[1]]
  
  if(length(parts) >= 2) {
    first <- parts[1]
    last <- parts[length(parts)]
    
    # Looking for matches in the People table
    match <- People %>%
      filter(tolower(nameFirst) == tolower(first),
             tolower(nameLast) == tolower(last))
    
    if(nrow(match) > 0) {
      return(match$playerID[1])
    }
  }
  return(NA)
}

# Getting IDs for everyone
voting_data$playerID <- sapply(voting_data$Name, get_playerID)

# Check if anyone didn't match
missing <- voting_data %>% filter(is.na(playerID))
if(nrow(missing) > 0) {
  cat("Couldn't find IDs for:", missing$Name, "\n")
}
```

## Making the final dataset

```{r}
# Creating the dataset in the same format as HallOfFame
HallOfFame_2025 <- data.frame(
  playerID = voting_data$playerID,
  yearID = 2025,
  votedBy = "BBWAA",
  ballots = total_ballots,
  needed = votes_needed,
  votes = voting_data$Votes,
  inducted = voting_data$inducted,
  category = "Player",
  needed_note = NA
)

# Removing anyone without a playerID
HallOfFame_2025 <- HallOfFame_2025 %>% filter(!is.na(playerID))

head(HallOfFame_2025, 10)
```

## Checking if it matches Lahman

```{r}
# Seeing if 2025 data is already in the Lahman database
lahman_2025 <- HallOfFame %>% filter(yearID == 2025)

if(nrow(lahman_2025) > 0) {
  cat("Found 2025 data in Lahman - comparing now...\n")
  cat("Lahman has", nrow(lahman_2025), "records\n")
  cat("I scraped", nrow(HallOfFame_2025), "records\n")
} else {
  cat("No 2025 data in Lahman yet\n")
}
```

## Saving the data

```{r}
write_csv(HallOfFame_2025, "HallOfFame_2025.csv")
cat("Saved to HallOfFame_2025.csv\n")
cat("Inducted:", sum(HallOfFame_2025$inducted == "Y"), "\n")
cat("Not inducted:", sum(HallOfFame_2025$inducted == "N"), "\n")
```

## Making the plot

```{r}
HallOfFame %>%
  ggplot(aes(x = yearID, fill = inducted)) +
  geom_bar() +
  xlim(c(1936, 2025)) +
  labs(title = "Hall of Fame Voting Over Time",
       x = "Year",
       y = "Count") +
  theme_minimal()
```

## My notes

**What worked:**
The scraping part was pretty straightforward once I figured out rvest. parse_number was really helpful for getting rid of the % signs. Matching the column names from the first row was easier than I thought

**What was tricky:**
Getting the playerIDs to match was annoying - some names just wouldn't work. Had to remember to convert everything to numeric or it stayed as text. Making sure the inducted column was Y/N and not TRUE/FALSE

**Questions for my team:**
Did everyone get the same number of total ballots? Anyone have a better way to match the playerIDs?
